---
title: "10_Importing_Data"
output: html_notebook
---

# Import Libraries
`tidyverse` is a collection of packages for R designed with a consistent methodology.

```{r setup}
require(tidyverse)
```

# Import Dataset
`readr` is used to import the datasets. Take note of the datetime format and the factor levels for events
```{r import data files}
driver_ids <- read_csv("data/driver_ids.csv", 
    col_types = cols(driver_onboard_date = col_datetime(format = "%Y-%m-%d %H:%M:%S")))

ride_ids <- read_csv("data/ride_ids.csv", 
    col_types = cols(driver_id = col_character(), 
        ride_distance = col_double(), ride_duration = col_double(), 
        ride_id = col_character(), ride_prime_time = col_double()))

ride_timestamps <- read_csv("data/ride_timestamps.csv", 
    col_types = cols(event = col_factor(levels = c("requested_at", 
        "accepted_at", "arrived_at", "picked_up_at", 
        "dropped_off_at")), ride_id = col_character(), 
        timestamp = col_datetime(format = "%Y-%m-%d %H:%M:%S")))
```

# Preview Data
```{r}
driver_ids
```

There are 937 drivers.

```{r}
ride_ids
```

There are 193,502 rides.

```{r}
ride_timestamps
```
looking at na timestamps:
```{r}
ride_timestamps %>% filter(is.na(timestamp))
```
filtering out na timestamps
```{r}
ride_timestamps <- ride_timestamps %>% filter(!is.na(timestamp))
dim(ride_timestamps)
```


```{r}
library(lubridate)
last_timestamp <- ride_timestamps %>% drop_na()
last_recorded_timestamp <- (max(last_timestamp$timestamp))
last_recorded_timestamp

first_recorded_timestamp <- (min(last_timestamp$timestamp))
first_recorded_timestamp

length_of_study <- last_recorded_timestamp-first_recorded_timestamp
length_of_study

```

# Combine Data into full dataset

```{r}
dim(full_join(ride_timestamps, ride_ids, by="ride_id"))
```
```{r}
dim(full_join(ride_timestamps, ride_ids, by="ride_id") %>% drop_na())
```

```{r}
979087-924094
# 8683 #na timestamp (rides in 2nd frame that you dont have time info for)
# 46310 #na driver id (rides in 3rd frame that you don't have driver/ridedistance/duration/primetime)
- 8683 - 46310
```

```{r}
58126 - 54993
```


```{r}
full_join(ride_timestamps, ride_ids, by="ride_id") %>% 
    filter(is.na(driver_id))
```
```{r}
ride_ids %>% 
    filter(ride_id == "0000d9b24d8ccdd991b76258e616fa01")
```
This shows that there are some rides in the `ride_timestamps.csv` that when searched in `ride_ids.csv` do not show up. Lyft did not give us a complete set of rides in the `ride_ids.csv` file. 

```{r}
full <- full_join(ride_ids,ride_timestamps,by="ride_id") %>%  full_join(driver_ids,by="driver_id")
full
```


```{r}
full <- full %>% drop_na()
full
```

Number of rows that were dropped:

```{r}
979170-921044 
```


## Calculating Ride Fare
```{r}

#defining constants
MILES.PER.METER <- 0.000621371
SECONDS.PER.MINUTE <- 60
USD.PER.MILE <- 1.15
USD.PER.MINUTE <- 0.22
USD.SERVICE.FEE <- 1.75
USD.BASE.FARE <- 2
MIN.FARE <- 5
MAX.FARE <- 400

earnings <-
    full %>% filter(event == "dropped_off_at") %>% select(ride_id, ride_distance, ride_duration, ride_prime_time) %>% mutate(unbound_driver_earnings = 
                                                                                                                                 (
                                                                                                                                     (
        USD.BASE.FARE + 
        USD.PER.MILE * ride_distance * MILES.PER.METER +
        0.22 * (ride_duration /SECONDS.PER.MINUTE)
                                                                                                                                     )
        * ((100 + ride_prime_time) / 100) + 
            USD.SERVICE.FEE 
                                                                                                                                 )
    ) %>% mutate(driver_earnings = ifelse(
        unbound_driver_earnings < MIN.FARE,
        MIN.FARE,
        ifelse(unbound_driver_earnings > MAX.FARE, MAX.FARE, unbound_driver_earnings)
    )) %>% select(ride_id, driver_earnings)

earnings
```

```{r}
full <- full_join(full, earnings, by="ride_id")
dim(full)
```

```{r}
write_csv(full, path = "data/full.csv")
```


