---
title: "40_Driver_Classification"
output: html_notebook
---
# Import Data
```{r setup}
library(tidyverse)

driver_attributes <- read_csv("data/driver_attributes_complete.csv")

driver_attributes
```


 
# Principle Component Analysis 


```{r}

driver_attributes <- driver_attributes %>% remove_rownames %>% column_to_rownames(var="driver_id") %>% select(-quit_within_90_days)

library("factoextra")
library("FactoMineR")
driver.pca <- PCA(driver_attributes, graph = FALSE)

```

```{r}
print(driver.pca)
```

## Scree plot

```{r}
eigenvalues <- driver.pca$eig
head(eigenvalues[, 1:2])
fviz_screeplot(driver.pca, ncp=10)
```

## Plot the correlations/loadings of the variables with the components

```{r}
head(driver.pca$var$coord)

fviz_pca_var(driver.pca)

```

```{r}
fviz_pca_contrib(driver.pca, choice = "var", axes = 1)
fviz_pca_contrib(driver.pca, choice = "var", axes = 2)
fviz_pca_contrib(driver.pca, choice = "var", axes = 1:2)
```

```{r}
fviz_pca_var(driver.pca, col.var="contrib")
```

```{r}
fviz_pca_ind(driver.pca, label="none")
```

# mixed clustering
Based off of: https://www.r-bloggers.com/clustering-mixed-data-types-in-r/
```{r setup2}
library(Rtsne) # for t-SNE plot
library(ggplot2) # for visualization

set.seed(1680) # for reproducibility

```

```{r}
driver_attributes <- read_csv("data/driver_attributes_complete.csv")

driver_attributes <- driver_attributes %>% remove_rownames %>% column_to_rownames(var="driver_id")
```


```{r}
gower_dist <- daisy(driver_attributes,
                    metric = "gower",
                    type = list(logratio = 3))
```

```{r}
summary(gower_dist)
```

```{r}
gower_mat <- as.matrix(gower_dist)

#most similar pair
driver_attributes[
  which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]),
        arr.ind = TRUE)[1, ], ]

#most dissimilar pair
driver_attributes[
  which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]),
        arr.ind = TRUE)[1, ], ]
```

```{r}
# Calculate silhouette width for many k using PAM

sil_width <- c(NA)

for(i in 2:10){
  
  pam_fit <- pam(gower_dist,
                 diss = TRUE,
                 k = i)
  
  sil_width[i] <- pam_fit$silinfo$avg.width
  
}

# Plot sihouette width (higher is better)

plot(1:10, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width)
```

plot peaks at 2, so 2 clusters are recommended

```{r}
pam_fit <- pam(gower_dist, diss = TRUE, k = 2)

pam_results <- driver_attributes %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results$the_summary
```
descriptive statistics for each cluster ^


these are the most representative samples of their clusters:
```{r}
driver_attributes[pam_fit$medoids, ]
```

```{r}
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)

tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))

ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))
```

this makes it look like there should be four clusters

now i will re run with four clusters


```{r}
pam_fit_4 <- pam(gower_dist, diss = TRUE, k = 4)

pam_results_4 <- driver_attributes %>%
  mutate(cluster = pam_fit_4$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results_4$the_summary
```
descriptive statistics for each cluster ^

these are the most representative samples of their clusters:
```{r}
driver_attributes[pam_fit_4$medoids, ]
```

looks like quit status and number of rides given are two major drivers of clustering

```{r}
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)

tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit_4$clustering))

ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))
```

looks like four clusters with a bit of messiness between cluster 1 and 2
