---
title: "60_Analyzing_Clusters"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

```{r}
full <- read_csv("data/full.csv", col_types = cols(timestamp = col_datetime(format = "%Y-%m-%d%.%H:%M:%S%.")))
full

driver_attributes <- read_csv("data/driver_attributes_complete.csv")
driver_attributes

cluster_assignments <- read_csv("data/cluster_assignments.csv", 
    col_types = cols(cluster = col_factor(levels = c("1", 
        "2", "3", "4"))))

```

```{r}
driver_attributes <- left_join(driver_attributes,cluster_assignments,by="driver_id")
driver_attributes
```

```{r}
cluster_names <- c("High-volume Quitters","Low-volume Quitters","Hustlers","Long-term Part-timers")
```


```{r}
summary_clusters <- driver_attributes %>% 
  group_by(cluster) %>% 
  summarise(
    `Average Earnings per Ride (USD)` = round(mean(average_earnings_per_ride), 2),
    
    `Average Ride Distance (miles)` = round(mean(average_ride_distance) * 0.000621371, 2),
    
    `Average Ride Duration (mins)` = round(mean(average_ride_duration) / 60 , 2),
    
    `Average Eagerness` = round(mean(scaled_average_eagerness), 2),
    
    `Average Responsiveness` = round(mean(scaled_average_response_time_ranking), 2),
    
    `Average Number of Rides Given` = round(mean(number_of_rides_given), 2),
    
    `Average Prime Time Weighted by Ride Duration` = round(mean(average_prime_time_weighted_by_ride_duration), 2),
    
    `Size`= n(),
    
    `Number of Days Worked` = round(mean(number_of_days_worked),2),
    
    `Average Number of Rides Per Day Worked` = round(mean(rides_per_day_worked),2)
  ) %>% 
  cbind(cluster_names) %>% 
  select(`Cluster Name` = cluster_names, 2:11)
  
summary_clusters
```

```{r}
write_csv(summary_clusters, path = "data/summary_clusters.csv")
```

# T-tests between the Husters and the Part-timers

Trying to see which variabes are statistically significant between the two non-quitter clusters

## average earnings

```{r}
variable <- "earnings"
earnings <- driver_attributes %>% 
  filter(cluster == 3 | cluster == 4) %>% 
  group_by(cluster) %>% 
  select(average_earnings_per_ride) %>% 
  group_split()

p <- t.test(earnings[[1]]$average_earnings_per_ride,earnings[[2]]$average_earnings_per_ride)$p.value 

if(p < 0.05) {
  print(paste0(variable, " is significant"))
}
```

## ride distance

```{r}
variable <- "ride distance"

distance <- driver_attributes %>% 
  filter(cluster == 3 | cluster == 4) %>% 
  group_by(cluster) %>% 
  select(average_ride_distance) %>% 
  group_split()

p <- t.test(distance[[1]]$average_ride_distance,distance[[2]]$average_ride_distance)$p.value 

if(p < 0.05) {
  print(paste0(variable, " is significant"))
}
```


## ride duration

```{r}
variable <- "ride duration"

duration <- driver_attributes %>% 
  filter(cluster == 3 | cluster == 4) %>% 
  group_by(cluster) %>% 
  select(average_ride_duration) %>% 
  group_split()

p <- t.test(duration[[1]]$average_ride_duration,duration[[2]]$average_ride_duration)$p.value

if(p < 0.05) {
  print(paste0(variable, " is significant"))
} else {
  print(paste0(variable, " is NOT significant"))
}
```

## number of rides

```{r}
variable <- "number of rides"

num <- driver_attributes %>% 
  filter(cluster == 3 | cluster == 4) %>% 
  group_by(cluster) %>% 
  select(number_of_rides_given) %>% 
  group_split()

p <- t.test(num[[1]]$number_of_rides_given,num[[2]]$number_of_rides_given)$p.value

if(p < 0.05) {
  print(paste0(variable, " is significant"))
} else {
  print(paste0(variable, " is NOT significant"))
}
```

## rides per day

```{r}
variable <- "number of rides per day worked"

num <- driver_attributes %>% 
  filter(cluster == 3 | cluster == 4) %>% 
  group_by(cluster) %>% 
  select(rides_per_day_worked) %>% 
  group_split()

p <- t.test(num[[1]]$rides_per_day_worked,num[[2]]$rides_per_day_worked)$p.value

if(p < 0.05) {
  print(paste0(variable, " is significant"))
} else {
  print(paste0(variable, " is NOT significant"))
}
```

# Driver Distribution
```{r}
driver_attributes %>% 
  ggplot(aes(x="", fill=cluster)) + geom_bar(width = 1) + coord_polar("y", start=0) +
  theme_minimal()
```

# Average Earnings Per Day for Non quitters
```{r}
full %>% 
  left_join(driver_attributes, by = "driver_id") %>% 
  filter(quit_within_90_days == F) %>% 
  select(driver_id, ride_id, driver_earnings) %>% 
  distinct() %>% 
  group_by(driver_id) %>% 
  summarise(total_earnings = sum(driver_earnings)) %>% 
  summarise(mean(total_earnings))
```

```{r}
3632.63/90
```


#Average Earnings per day in career for quitters
```{r}
driver_attributes %>% 
  filter(quit_within_90_days == T) %>% 
  select(driver_id,career_length_in_days,average_earnings_per_ride, number_of_rides_given, cluster) %>% 
  mutate(total_earnings = number_of_rides_given * average_earnings_per_ride) %>% 
  mutate(average_earnings_per_day_in_career = total_earnings/career_length_in_days) %>% 
  group_by(cluster) %>% 
  summarise(mean(average_earnings_per_day_in_career))
```

# Average 90 day earnings by cluster (for nonquitters)

```{r}
driver_attributes %>% 
  filter(quit_within_90_days == F) %>% 
  select(driver_id,average_earnings_per_ride, number_of_rides_given, cluster) %>% 
  mutate(total_earnings = number_of_rides_given * average_earnings_per_ride) %>% 
  group_by(cluster) %>% 
  summarise(`three-month-total` = mean(total_earnings)) %>% 
  mutate(daily_earnings = `three-month-total` / 90)
```
